---
title: "An Simple Introduction to R"
output: pdf_document
date: "2023-01-16"
author: Keyu Long
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic Data Types

## Assign Value to Variables

```{r}
x = 4
y <- 2
# both work
```

## Numeric Operation

```{r}
x + y 
x - y
x * y
x / y
x ^ y
```

## Characteristic

```{r}
z = 'math'
z = "math" # both are the same
u = '1'
```

## Logical

```{r}
a = TRUE
b = T
c = F
```

## Special

```{r}
w = NA # not a number / represnt NONE in R
```

## Class

```{r}
# used to see the type of variable
class(x)
class(z)
class(u)
class(w)
```

### Type Conversion

```{r}
v = as.numeric(u) # convert from a char to number, can only convert reasonable values
class(v)
is.numeric(v) 
```

# Logical Operations

## Simple Comparisons

```{r}
x == 4 # equal to

x != 1 # not equal to

!(x == 4) # negation of equal

x >= 1
y < 1
```

## Multiple Logical Operations

```{r}
x > 3 & y > 1 # and 
x > 3 | y < 0 # or
```

# Data Structure

## Vectors

```{r}
# define vectors
vec1 = c(45, 43, 46, 48, 51, 46, 50, 47, 46, 45)
vec2 = 1:10 # R does not support syntax like 1:20:2, rather use seq()
vec3 = seq(1, 20, 2)

# to seek for help
help('seq')

vec4 = c('math 170A', 'math 170B', 'math 170C', 'math 180A', 'math 180B', 'math 180C', 'math 181A', 'math 181B', 'math 183', 'math 185')
```

### Operations

```{r}
# Basic Functions
sum(vec2)
mean(vec1)
median(vec1)
max(vec1)
min(vec1)
var(vec1)
summary(vec1) # summary returns all of the above info
```

```{r}
# standard deviation
sd(vec1)

# length of a vector
length(vec4)
```

```{r}
# indexing of vector

# get the i-th value
vec[2]
vec1[c(1, 3, 5)] # get the 1st, 3rd, 5th element at the same time

# find the index
which(vec1 == 45)
which(vec1 == max(vec1)) # find the index of the max value
```

### Sort

```{r}
order(vec1, decreasing = T) # return ordered index
```

### Naming

```{r}
# name the value of each element
vec5 = c(90, 91, 95, 96, 98, 85, 93, 95, 92, 88)
names(vec5) <- vec4
vec5
```

### Between Vectors

```{r}
# direct operation between vectors are element-wised
vec1 + vec2
vec1 - vec3
vec1 * vec2 # Hadamard product not Dot product
as.numeric(vec1 %*% vec2) # dot product and get the element out
```

## Matrices

```{r}
help('matrix')
```

```{r}
# declare new matrices
A = matrix(1:6, nrow = 2, ncol = 3) # default: by column
B = matrix(1:6, nrow = 2, ncol = 3, byrow = T)
C = matrix(2:7, nrow = 2, ncol = 3)
```

### Operations

```{r}
# Basic functions
t(A) # transpose
dim(A) # dimension of a matrix

# Again, direct operations are element-wised
A*C # Hadamard product

# dot product
A %*% t(C)
```

## Table

```{r}
sample_data1 = rpois(100, 5)
sample_data2 = rpois(100, 2)
ts = table(sample_data1) # count frequency
ts
sample_data1
```

# Packages Management

```{r}
# install packages
install.packages('pracma')
# load pakcages
library(pracma)

# then we can use product method
dot(vec1, vec2)

# unload / detach
detach('package:pracma')
```

# Data Frame

```{r}
library(MASS)
dataset = Cars93 # specific dataset
class(dataset)
```

```{r}
# check the strcuture of a dataframe
str(dataset)
```

```{r}
# summary statistics
summary(dataset)
```

## Operations

```{r}
# visit one variable
dataset$Weight
```

```{r}
#extract a varible
weight <- dataset$Weight
mpg_city <- dataset$MPG.city
```

```{r}
#attach a dataset / directly load it
attach(Cars93)
Weight
MPG.city
```

```{r}
# detach the dataset
detach(Cars93)
```

# Plots

### Histogram

```{r}
# for numeric variables
hist(dataset$Weight)
dev.off() # clear all plots
```

```{r}
# add titles
hist(dataset$Weight, main = "Histogram of Car Weights", xlab = 'Weight', col = 'grey')
```

### Barplot

```{r}
# for categorical variables
manu <- table(dataset$Manufacturer)
JP_cars <- which(names(manu) %in% c('Honda', 'Acura', 'Toyota', 'Lexus', 'Nissan', 'Infiniti'))
barplot(manu[JP_cars], main = 'Japenese Cars',
        xlab = 'Manufacturers', ylab = 'Frequency', axes = T, axis.lty = 1)
```

### Boxplot

```{r}
boxplot(mpg_city)
boxplot(mpg_city, xlab = "MPG in City", col = 'yellow', outpch = 8, outcol = 'red')
boxplot(mpg_city, xlab = "MPG in City", col = 'yellow', outpch = '*', outcol = 'red')
```

### Scatter Plot

```{r}
plot(weight, mpg_city)
plot(weight, mpg_city, main = "relation btw car weight & mpg in city",
     xlab = 'car weight', ylab = 'mpg in city', pch = 19, cex = 0.8, col = 'darkgreen')
```

# Linear Regression

```{r}
# draw the scatter first
plot(weight, mpg_city, main = "relation btw car weight & mpg in city",
     xlab = 'car weight', ylab = 'mpg in city', pch = 19, cex = 0.8, col = 'darkgreen')

lin <- lm(formula = mpg_city ~ weight)
abline(lin, lwd = 1.5, col = 'red') # draw the fitting line
```
